## NCAA Football Attendance

Let's load the tidyverse.

```{r}
library(tidyverse)
```

And read in the attendance data:

```{r}
attendance <- read_csv("https://dwillis.github.io/sports-data-files/attendance.csv")
```

Using the tidyverse's `pivot_longer` function, turn the currently wide data into longer data and save it as a new dataframe called `attendance_longer`:

```{r}
attendance_longer <- attendance |> pivot_longer(cols=starts_with("20"), names_to="Year", values_to="Attendance")
```

Let's try to figure out whether football attendance has returned to "normal" by 2022, defining "normal" as 2019 figures. For that we'll need to isolate those two years using `filter()` and get rid of any NAs:

```{r}
attendance_two_years <- attendance_longer |> filter(Year == 2019 | Year == 2022) |> na.omit()
```

Now, in order to calculate a percentage change between those two years, we'll need to go back to wider data and again ditch teams where we don't have both years:

```{r}
attendance_wide <- attendance_two_years |> pivot_wider(names_from=Year, values_from=Attendance) |> na.omit()
```

Finally, we'll add a percentage change column to that dataframe (remember, it's (New-Old)/Old) and arrange it to see the results.

```{r}
attendance_wide <- attendance_wide |> mutate(percent_change = ((`2022`-`2019`)/`2019`)*100) |> filter(percent_change < 2 & percent_change > -2) |> arrange(desc(percent_change))
```

Some of these clearly are influenced by performance on the field. Below, describe the most interesting results (positive or negative) and what you think might be going on with them. Feel free to do some googling and describe what other data would be helpful to clarify what the best stories here might be. What surprises you about these results?

**ANSWER**
Hawaii had the biggest drop off, almost -70%. In 2019, Hawaii was 10-5, but in 2022, they were 3-10. Conversely, Arkansas had a ~43% increase in attendance, with a 2019 record of 2-10 (0-8 in SEC games!) but a 7-6 record in 2022, indicating some sort of an uptick in the quality of the program.

Interestingly, Miami and Oklahoma (along with 8 other schools), had a less than 1% fluctuation (either up or down) on their numbers. 7 of these ten schools were P5 schools, indicating strong, consistent attendance figures. Shoutout New Mexico State (probably the smallest of these schools) for staying consistent!


