```{r}
library(tidyverse)
```

```{r}
logs <- read_csv("../data/cbblogs1523.csv")
```

```{r}
logs <- logs |> mutate(
  Differential = TeamScore - OpponentScore, 
  NetRebounds = TeamTotalRebounds - OpponentTotalRebounds,
  TurnoverMargin = TeamTurnovers - OpponentTurnovers)

head(logs)
```

```{r}
rebounds <- lm(Differential ~ NetRebounds, data=logs)
summary(rebounds)
```

Low (good) p-value, a lower-than-good R-squared value, and the actual equation looks okay. Looks like rebounds might be a good predictor, but would be stronger when compared with other factors.

```{r}
model1 <- lm(Differential ~ NetRebounds + TurnoverMargin, data=logs)
summary(model1)
```

```{r}
library(Hmisc)
```

```{r}
simplelogs <- logs |> select_if(is.numeric) |> select(-Game) |> select(Differential, NetRebounds, TurnoverMargin, TeamFGPCT, TeamTotalRebounds, OpponentFGPCT, OpponentTotalRebounds)
```

```{r}
cormatrix <- rcorr(as.matrix(simplelogs))

cormatrix$r
```

```{r}
model2 <- lm(Differential ~ NetRebounds + TurnoverMargin + TeamFGPCT + OpponentFGPCT, data=logs)
summary(model2)
```

```{r}
logs |> 
  filter(
    Team == "Maryland" & Season == '2022-2023'
    ) |> 
  summarise(
    meanNetRebounds = mean(NetRebounds),
    meanTurnoverMargin = mean(TurnoverMargin),
    meanTeamFGPCT = mean(TeamFGPCT),
    meanOpponentFGPCT = mean(OpponentFGPCT)
  )
```

```{r}
logs |> 
     filter(
         Team == "Maryland" & Season == '2022-2023'
     ) |> summarise(avg_score = mean(TeamScore), avg_opp = mean(OpponentScore))
```


```{r}
library(glmnet)
newlogs <- na.omit(logs)
y <- newlogs$Differential
z <- newlogs |> select_if(is.numeric)
x <- data.matrix(subset(z, select = -c(Differential, TeamScore, OpponentScore)))
```

```{r}
cv_model <- cv.glmnet(x, y, alpha = 1)

best_lambda <- cv_model$lambda.min
best_lambda
plot(cv_model)
```

```{r}
best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
```

Using the Lasso Regression model with cross validation, we can see which factors are BEST to predict differential. I eliminated differential, as well as team and opponent score as factors, since including these factors would severely skew and zero out other factors. Based on this, there are a handful of factors that don't really matter. Lasso regressions are used to handle the issue of multicollinearity, and automatically handles factor selection by applying a penalty to selecting non-zero values for values for each factor, forcing the model to pick the best factors.

Best Factors now include TeamFG, TeamFGPCT, Team3P, Team3PPCT, TeamFT, TeamAssists, and the same factors (except for Opponent 3 point %), as well as NetRebounds and Turnover margin.

Intuitively, the selected factors make sense, as these factors most contribute to scoring, and therefore a differential at the end of games.

```{r}
best_model
```

Penalization parameter of 0.1325
13 Non-Zero Parameters
99.9% of deviance explained by regressors

########## Early Season Analysis, November 9, 2023

```{r}
suppressMessages(allgames <- read_csv("https://raw.githubusercontent.com/dwillis/dwillis.github.io/main/docs/sports-data-files/cbblogs1524.csv"))
suppressMessages(allumd <- allgames |> filter(TeamFull == "Maryland Terrapins"))
library(lubridate)
library(ggplot2)
```

```{r}
#str_detect(allumd$W_L, "W")
allumd <- allumd |>
  mutate(winloss = case_when(
    grepl("W", W_L) ~ 1, 
    grepl("L", W_L) ~ 0)
) 

umdSeasons <- allumd |> 
  group_by(Season) |> 
  summarise(
    wins = sum(winloss), 
    FG = sum(TeamFG), 
    FGA = sum(TeamFGA), 
    FGPercent = (sum(TeamFG)/sum(TeamFGA))
  )

umdEarlyGames <- allumd |> filter(Game <4)

umdEarlyGames <- umdEarlyGames |> 
  group_by(Season) |> 
  summarise(
    wins = sum(winloss), 
    FG = sum(TeamFG), 
    FGA = sum(TeamFGA), 
    FGPercent = (sum(TeamFG)/sum(TeamFGA))
  )
```

```{r}
ggplot() +
  geom_point(data=umdSeasons, aes(x=Season, y = FGPercent), color="black") +
  geom_point(data=umdEarlyGames, aes(x=Season, y = FGPercent), color="red") +
  geom_point(data=umdEarlyGames, aes(x=Season, y = wins/60), color="blue") +
  geom_point(data=umdSeasons, aes(x=Season, y = wins/60), color="green") +
  scale_y_continuous(name = "Field Goal Percentage, as a decimal",
    sec.axis = sec_axis( trans=~.*60, name="Wins")) +
  labs()
```
```{r}
windiff <- c(16, 21, 13, 12, 19, 0)
umdSeasons <- umdSeasons |> mutate(earlyFGPercent = umdEarlyGames$FGPercent,  winsafter3 = windiff, FGgap = umdSeasons$FGPercent - earlyFGPercent)
```

Trying to model FG differential in previous seasons from early games to the rest of the season and how many wins that nets you. (Do teams that cool off more win less, regardless of how hot they started?)

```{r}
ggplot() + 
  geom_point(data=umdSeasons, aes(x=FGgap, y=winsafter3))
```

