---
title: "Women's Soccer Data"
output: html_notebook
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1kTSrnl1URmcexIWLDMWqDnp3q-c2OS45uifnhjp10n4/edit#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2023 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:


```{r}
matches_2023 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2023.csv")
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2023)
glimpse(matches_2023)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1. How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal.

```{r}
matches_2023 |>
  group_by(team) |>
  summarise(
    shots = sum(sh_att),
    shotsOnGoal = sum(so_g),
    sOGPercent = (shotsOnGoal/shots)*100
  ) |> 
  filter(str_detect(team, "Maryland"))
```
Based on the above function, (I used string filtering to find teams that had "Maryland" in their name to avoid clicking through the 50+ pages), Maryland Women's Soccer has a shots on goal percentage of 53.57%. They have attempted a total of 56 shots, of which 30 have been on goal.

Resource used for question 1: https://stackoverflow.com/questions/22850026/filter-rows-which-contain-a-certain-string


2. What about the same shot stats from last season's first four games? You'll need to load the matches from 2022 - just change the season in the URL - and filter by date to find the first four games.

```{r}
matches_2022 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2022.csv")

matches_2022 |>
  filter(str_detect(team, "Maryland Terrapins")) |>
  filter(date < '2022-09-02') |>
  group_by(team) |>
  summarise(
    shots = sum(sh_att),
    shotsOnGoal = sum(so_g),
    sOGPercent = (shotsOnGoal/shots)*100
  )


```
It would appear that in the first 4 games of the 2022 season, Maryland Women's Soccer had 48 total shots, 23 of which were on target, for a shots on goal percentage of 47.92%



3. What about draws - Maryland started 2022 with 4 ties. Have any other teams had as many this season?

```{r}
matches_2023 |>
  filter(str_detect(outcome, "Draw")) |>
  group_by(team) |>
  summarise(
    draws = n()
  ) |>
  filter(draws == 4)
```
There are 3 teams in 2023 that have 4 draws so far: Denver, Indiana State, and Maryland themselves.



4. How many total corner kicks has Maryland had so far and how many have its opponents have? Calculate the difference between the two totals for every team.

```{r}
head(matches_2023)

matches_2023 |>
  group_by(team) |>
  summarise(
    cornersFor = sum(corners),
    cornersAgainst = sum(defensive_corners)
  ) #|>
  #ilter(str_detect(team, "Maryland Terrapins"))
```

So far this season, Maryland has taken 13 corners, and allowed 20. To find these values, we can remove the "#" characters to un-comment the code in lines 91 and 92.

## In class work for Thursday, September 14, 2023 below (SIGNIFICANCE TESTING):

Hypothesis: There is more scoring in Women's Soccer this season than last season.

```{r}
matches_2022 <- matches_2022 |> mutate(total_goals = goals + defensive_goals)

matches_2023 <- matches_2023 |> mutate(total_goals = goals + defensive_goals)
```
```{r}
t.test(matches_2022$total_goals, matches_2023$total_goals)
```

The results of the total goals t-test indicate that in over 99% of simulations, there is a statistically significant relationship between the two variables, so the rise in scoring is notable. Null hypothesis rejected.

Publication sentence: Is NCAA women's soccer more engaging this season? One might argue yes, seeing as there is a statistically significant increase in the average goals score per game!
```{r}
matches_2022_early <- matches_2022 |> filter(date < '2022-09-14')
matches_2022_early <- matches_2022_early |> mutate(total_goals = goals + defensive_goals)
```

```{r}
t.test(matches_2022_early$total_goals, matches_2023$total_goals)
```

P value is much higher, closer to the 0.05 rejection threshold. t-value is good, but it would indicate that the scoring in non-conference play is now closer to being "noisy" than having any sort of statistical significance.

Publication: The statistics show us that for most viewers, the competitive level in NCAA women's soccer non-conference play has maintained between last season and this one.