```{r}
library(tidyverse)
```


Questions from ELMS:

Does that data already exist? Can you get it?

The data already exists, and I've added it to the "TransferMarkt Data" Folder in my presentations folder.
```{r}
clubs <- read_csv("~/Desktop/TransferMarkt Data/clubs.csv")
players <- read_csv("~/Desktop/TransferMarkt Data/players.csv")
player_valuations<- read_csv("~/Desktop/TransferMarkt Data/player_valuations.csv")
games <- read_csv("~/Desktop/TransferMarkt Data/games.csv")
game_events <- read_csv("~/Desktop/TransferMarkt Data/game_events.csv")
comps <- read_csv("~/Desktop/TransferMarkt Data/competitions.csv")
club_games <- read_csv("~/Desktop/TransferMarkt Data/club_games.csv")
appearances <- read_csv("~/Desktop/TransferMarkt Data/appearances.csv")

head(game_events)
```
```{r}
club_games
```


Do you need assistance in getting it?
Nope! It's all easy to 


What time frame do you want/need data from?
The data I have goes back to the 2012-13 season and spans a wide range of leagues, so this amount of data is more than enough.

Will you have to improve or augment the data in order to answer your question? If so, how?
Yes, and a lot. It's structured as a relation DB, so there will need to be a lot of joins and zips to compress player stats. I'm planning on looking at discussions and posted
code on Kaggle to see if there's anything that will speed up my analysis process

Are there other examples of this kind of work (specific or general)? If so, provide them.

https://www.kaggle.com/code/luisgasparcordeiro/market-value-eda -> some inspiration from this, but it's nowhere near the depth I want to get to
https://www.kaggle.com/code/luisgasparcordeiro/player-stats-by-season -> python code to aggregate stats. using this, I did some googling and figured out how to translate
the python code that user luisgasparcordeiro made into R, which can be seen below.


```{r}
player_stats <-function(name, year) {

df <- merge(appearances, games, by="game_id", all.x=TRUE)

#nrow(df)
df$goals_for <- ifelse(df$home_club_id == df$player_club_id, df$home_club_goals,
                           ifelse(df$away_club_id == df$player_club_id, df$away_club_goals, NA))

df$goals_against <- ifelse(df$home_club_id == df$player_club_id, df$away_club_goals,
                               ifelse(df$away_club_id == df$player_club_id, df$home_club_goals, NA))
    df$clean_sheet <- ifelse(df$goals_against == 0, 1,
                            ifelse(df$goals_against > 0, 0, NA))
    

if (year == "all") {
  df <- df |> group_by(player_id, player_name, season) |> 
  summarise(
    g = sum(goals),
    a = sum(assists),
    cs = sum(clean_sheet),
  ) |> filter(player_name == name)
}
else {
  df <- df |> group_by(player_id, player_name, season) |> 
  summarise(
    g = sum(goals),
    a = sum(assists),
    cs = sum(clean_sheet),
  ) |> filter(player_name == name) |> filter(season == year)
}

df

}
 

```
```{r}
messi <- player_stats("Lionel Messi", "all")
#messivals <- merge(messi, player_valuations, by="player_id", all.x = TRUE)
messisvalue <- player_valuations |> filter(player_id == 28003)
library(ggplot2)
ggplot(messisvalue, aes(x=messisvalue$datetime, y=messisvalue$market_value_in_eur)) +
  geom_line()
```
```{r}
ggplot(messi, aes(x=messi$season, y=(messi$g +messi$a) )) +
  geom_line()
```



```{r}
player_stats('Cristiano Ronaldo', "all")
```


